\section*{Abstract} %% <= Insert Title: Abstract vs. Summary
\addcontentsline{toc}{section}{Abstract}
\markright{Abstract}

This report delivers an in-depth exploration of the Advanced Encryption Standard (AES), 
a cornerstone in contemporary symmetric block cipher design and a vital element in today’s cryptographic frameworks. 
It starts by placing AES within its historical backdrop, 
shedding light on the shortcomings of its predecessor, 
the Data Encryption Standard (DES), 
which AES was explicitly developed to address. 
The document proceeds to unpack the inner workings of AES—detailing the mechanisms of both encryption and decryption, 
the structure of key expansion, and the underlying mathematics based on operations in Galois Fields. 
Core algorithmic steps are carefully explained, 
including SubBytes, ShiftRows, MixColumns, and AddRoundKey. 
To support practical understanding, the report features a working AES-128 implementation written in Python, thoroughly tested against the official NIST-provided test vectors to ensure accuracy. 
The report further explores real-world implementations of AES across domains such as wireless communication, 
secure cloud storage, and encrypted messaging, emphasising its broad adoption. 
It also offers a balanced assessment of AES’s strengths, 
such as its strong security guarantees and computational efficiency, 
while acknowledging specific practical issues like key distribution challenges and the absence of native authentication features. 
In its final analysis, 
the report examines how AES stands up to emerging quantum computing threats, 
concluding that the algorithm remains a relevant and reliable choice for modern and future cybersecurity demands.
