\section{Introduction to Cryptography}

\subsection{Why Do We Need Encryption?}

With the development of technology, there are more ways of communication. 
Many years ago, people used Morse code; now, people write text messages on smartphones through apps that claim to provide end-to-end encryption. 
But where does this need come from, and why do we need encryption?

Through encryption, we secure our privacy, prevent unauthorized access to our lives, protect government systems against attacks, and, for many companies, keep corporate secrets safe from competitors. 
Unauthorized listening to private communication is known as \textit{eavesdropping}.

\subsection{Symmetric and Asymmetric Encryption}

Symmetric encryption uses one key to encrypt and decrypt data. 
Before communication, the sender and receiver must securely share the key.

Asymmetric encryption uses a pair of keys: a public key and a private key. 
The public key is shared with everyone, while the private key is unique to the receiver. 
Thus, the public key encrypts, and the private key decrypts.

AES and DES are examples of symmetric encryption, whereas RSA is an example of asymmetric encryption.

The main difference between symmetric and asymmetric encryption is that symmetric encryption is faster since it uses the same key for both encryption and decryption. 
Asymmetric encryption is slower but solves the key exchange problem. Therefore, they are often combined: asymmetric encryption securely shares a key, and symmetric encryption is used for fast communication.